Sulfur Dioxide on Water's Surface

### 2011-02-24 EquilibratedOrientation.tex
resized images to fit in the PDF
not sure

The location of \suldiox~within the water surface is slightly different between the two equilibrium systems. The single \suldiox~molecule is generally located just below the 0\angs position of the surface, and the distribution of the concentrated system \suldiox~molecules is centered above the surface. This indicates that in both systems the \suldiox~is highly surface active, but the locations of the peak \suldiox~densities differ. The single molecule spends most of its time bound within the top water monolayer, and often traversing below that location by several\angs. The saturated system's surface molecules sit on top of the monolayer. The angular distribution of the single \suldiox~is concentrated primarily in $\cos(\theta)>0$ taking an orientation of the bisector pointing out of the water surface. This same distribution occurs in the saturated system for positions below 0\angs, but the orientation becomes isotropic for molecules located above 0\angs, on top of the water surface.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2orientationsmall.png}
		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface. Distributions are shown for $\cos(\theta)$ (left column) and $\cos(\phi)$ (right column) for both the neat-\wat~(top) and high \suldiox~concentration (bottom) systems. For both systems the $\theta$ distributions suggest \suldiox~bound to the water surface with the sulfur pointing towards the water slab, and the oxygens pointing to the gas phase. In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry.}
		\label{fig:so2-orientation}
	\end{center}
\end{figure}

>>>>

The location of \suldiox~within the water surface is slightly different between the two equilibrium MD systems (see figure \ref{fig:density}). The single \suldiox~molecule is generally located just below the 0\angs position of the surface, and the distribution of the concentrated system \suldiox~molecules is centered above the surface location. This indicates that in both systems the \suldiox~is highly surface active, but the locations of the peak \suldiox~densities differ. The single molecule spends most of its time bound within the top water monolayer, and often traversing below that location by only a few\angs. The saturated system's surface molecules sit on top of the monolayer. The angular distribution of the single \suldiox~is concentrated primarily in $\cos(\theta)>0$, such that the bisector points out of the water surface. This same distribution occurs in the saturated system for positions below 0\angs, but the orientation becomes isotropic for molecules located above 0\angs, on top of the water surface. Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2-angles/so2-angles.png}
		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface. Distributions are shown for $\cos(\theta)$ (left column) and $\cos(\phi)$ (right column) for both the neat-\wat~(top) and high \suldiox~concentration (bottom) systems. For both systems the $\theta$ distributions suggest \suldiox~bound to the water surface with the sulfur pointing towards the water slab, and the oxygens pointing to the gas phase. In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry.}
		\label{fig:so2-orientation}
	\end{center}
\end{figure}

### 2011-02-24 Methods.tex
not sure

% need to add in the specifics about the SMD routine, force constants, calculations, etc.
A second set of simulations involved preparing an equilibrated water slab as in the surface equilibrated method above. However, for both the neat-water and saturated systems, a single \suldiox~was introduced 20\angs above the water slab surface, with the sulfur atom tethered to its initial position. The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations. Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the low and high concentrations) to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.

Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-\wat~slab and the saturated-\wat~slab configurations before steering the \suldiox~towards the water bulk.

>>>>

% need to add in the specifics about the SMD routine, force constants, calculations, etc.
A second set of simulations involved an equilibrated water slab as in the surface equilibrated method above. However, for both the neat-water and saturated systems, a single \suldiox~was introduced 20\angs above the water slab surface, with the sulfur atom tethered to its initial position. The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations. Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the low and high concentrations) to guide the \suldiox~down towards a tethered water-oxygen 15\angs under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab. Each of the SMD simulations were performed for a total of 200 ps, using a timestep of 1 fs, and taking snapshots of the system every 25 fs. Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-water slab and the saturated-\wat~slab configurations before steering the \suldiox~towards the water bulk.

A separate set of SMD simulations were performed similarly, but with the tethering to one of the \suldiox-oxygens. This was done to ensure that the orientation of the \suldiox~during the adsorption transit was not an artifact of the choice of atom used for tethering. The simulations produced the same results (not shown) for the orientational analyses, so the data from the original tethering scheme was used.

### 2011-02-24 TransitOrientation.tex
Rearranging the filenames and working on the h2o-orientation results
resized images to fit in the PDF
not sure

\section{Steered MD}

@@

	The orientation of a \suldiox~molecule adsorbing to an aqueous surface was determined through an analysis process to find the angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}). For each timestep of the SMD simulations the orientation angles were determined for a single \suldiox~as it was pulled into the water slab, both in the neat-\wat~and saturated slab systems. The resulting values were collected for the 50 simulations for both systems, and the angle values were binned indexed with the position of the \suldiox~relative to the water slab surface location. The resulting 2-dimensional histograms shown in figure \ref{fig:so2-transit-angles} show the orientation trend for the \suldiox~in the gas phase, near the aqueous surface, and after entering the water bulk.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2transitorientationsmall.png}
		\caption{}
		\label{fig:so2-transit-angles}
	\end{center}
\end{figure}

	From its starting position 20\angs above the water surface, until the \suldiox~moves to within 5\angs of the surface of both systems, the orientation is isotropic in both $\theta$ and $\phi$. Upon approaching the saturated surface at approx. 5\angs, the \suldiox~$\phi$ orientation distribution begins to cluster near $\cos(\phi)=1$, with the $\theta$ distribution remaining more isotropic, but with a trend of $\cos(\theta)>0$. At 5\angs the \suldiox~begins interacting with the other \suldiox~molecules that sit atop the water, and the resulting orientation distribution suggests that contact with the \suldiox~layer causes the adsorbing molecule to lie more flat to the surface. Within this same region from 0-5\angs above the surface, the $\phi$ distribution of the neat-\wat~system remains isotropic. The intensity of the distribution (darker blue) also indicates that the \suldiox~spends little time in this region, as it is quickly adsorbed further into the water bulk.

	Just below the water surface as the distance moves into negative values, both systems show clear orientational preferences. In both systems the bisector orientation clusters such that $\cos(\theta)<0.5$, with most of the distribution intensity around $\cos(\theta)=1$. This suggests that \suldiox~in a water surface orients with the sulfur pointing into the water bulk, and the oxygens pointing towards the surface water molecules.

	Differences occurr between the system orientation profiles once the \suldiox~has moved further than 10\angs below the surface. While both $\theta$ and $\phi$ hold the same trend from just under the surface through to the bulk of the slab in the saturated system, the orientation profile becomes much more isotropic past 10\angs under the surface of the neat-\wat~system. The orientation effect is shallower under the neat-\wat~surface. However, the orientation distribution peaks most intensely, and more tightly clustered with the \suldiox~bisector aligning with the surface normal, just underneath the neat-\wat~surface than the saturated one. The distributions thus show an orienting force extending deeper into the surface of the saturated system than the neat-\wat~system. In terms of how deeply the \suldiox~molecule is oriented, the interfacial region of the saturated system is larger and extends further into the aqueous bulk.

>>>>

\section{Steered MD Transit Simulations}

@@

	The orientation of a \suldiox~molecule adsorbing to an aqueous surface was determined through an analysis process to find the angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}). For each timestep of the SMD simulations the orientation angles were determined for the transiting \suldiox~as it was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems. The angle cosine values were collected for the 50 simulations for both systems, and the binned with the position of the \suldiox~relative to the water slab surface location. The resulting 2-dimensional histograms shown in figure \ref{fig:so2-transit-angles} show the orientation trend for the \suldiox~in the gas phase (above 0\angs), near the aqueous surface (near 0\angs), and after entering the water bulk (below 0\angs). Only the surface into which the transiting \suldiox~was steered is shown for the saturated system.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2-angles/so2-angles-transit.png}
		\caption{\suldiox~molecular orientation distributions during SMD transit simulations into an aqueous slab. Both the neat-water (top row) and saturated (bottom row) data sets were analyzed to determine the angles $\theta$ (left column), and $\phi$ (right column) of the transiting \suldiox. The distributions show angle cosines plotted against the distance of the \suldiox~to the location of the water surface. The data sets were averaged at each distance for the respective 50 simulations.}
		\label{fig:so2-transit-angles}
	\end{center}
\end{figure}

	%From its starting position 20\angs above the water surface, until the \suldiox~moves to within 5\angs of the surface of both systems, the orientation is isotropic in both $\theta$ and $\phi$. Upon approaching the saturated surface at approx. 5\angs, the \suldiox~$\phi$ orientation distribution begins to cluster near $\cos(\phi)=1$, with the $\theta$ distribution remaining more isotropic, but with a trend of $\cos(\theta)>0$. At 5\angs the \suldiox~begins interacting with the other \suldiox~molecules that sit atop the water, and the resulting orientation distribution suggests that contact with the \suldiox~layer causes the adsorbing molecule to lie more flat to the surface. Within this same region from 0-5\angs above the surface, the $\phi$ distribution of the neat-\wat~system remains isotropic. The intensity of the distribution (darker blue) also indicates that the \suldiox~spends little time in this region, as it is quickly adsorbed further into the water bulk.

	%Just below the water surface as the distance moves into negative values, both systems show clear orientational preferences. In both systems the bisector orientation clusters such that $\cos(\theta)<0.5$, with most of the distribution intensity around $\cos(\theta)=1$. This suggests that \suldiox~in a water surface orients with the sulfur pointing into the water bulk, and the oxygens pointing towards the surface water molecules.

	%Differences occurr between the system orientation profiles once the \suldiox~has moved further than 10\angs below the surface. While both $\theta$ and $\phi$ hold the same trend from just under the surface through to the bulk of the slab in the saturated system, the orientation profile becomes much more isotropic past 10\angs under the surface of the neat-\wat~system. The orientation effect is shallower under the neat-\wat~surface. However, the orientation distribution peaks most intensely, and more tightly clustered with the \suldiox~bisector aligning with the surface normal, just underneath the neat-\wat~surface than the saturated one. The distributions thus show an orienting force extending deeper into the surface of the saturated system than the neat-\wat~system. In terms of how deeply the \suldiox~molecule is oriented, the interfacial region of the saturated system is larger and extends further into the aqueous bulk.
